---
title: "generate_puzzle()"
subtitle: "The main entry point for creating puzzles"
---

```{r}
#| include: false
source("quarto/_setup.R")
```

## Description

`generate_puzzle()` is the unified entry point for creating all puzzle types in jigsawR. It handles rectangular, hexagonal, concentric, voronoi, and random puzzles through a single consistent interface.

## Usage

```r
generate_puzzle(
  type = "rectangular",
  grid,
  size,
  seed = NULL,
  offset = 0,
  layout = "grid",
  fusion_groups = NULL,
  save_files = FALSE,
  output_dir = "output",
  ...
)
```

## Arguments

### Required Parameters

| Parameter | Type | Description |
|-----------|------|-------------|
| `type` | string | Puzzle type: `"rectangular"`, `"hexagonal"`, `"concentric"`, `"voronoi"`, `"random"` |
| `grid` | numeric vector | Grid dimensions (interpretation depends on type) |
| `size` | numeric vector | Physical dimensions in mm |

### Grid Parameter by Type

| Type | Grid Format | Example |
|------|-------------|---------|
| rectangular | `c(cols, rows)` | `c(4, 3)` = 4x3 grid |
| hexagonal | `c(rings)` | `c(3)` = 3 rings |
| concentric | `c(rings)` | `c(4)` = 4 rings |
| voronoi | `c(n_cells)` | `c(25)` = 25 cells |
| random | `c(n_pieces)` | `c(20)` = 20 pieces |

### Size Parameter by Type

| Type | Size Format | Example |
|------|-------------|---------|
| rectangular | `c(width, height)` | `c(400, 300)` |
| hexagonal | `c(diameter)` | `c(250)` |
| concentric | `c(diameter)` | `c(200)` |
| voronoi | `c(width, height)` | `c(400, 300)` |
| random | `c(width, height)` | `c(400, 300)` |

### Optional Parameters

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 | Piece separation distance in mm |
| `layout` | string | "grid" | Layout algorithm: `"grid"` or `"repel"` |
| `fusion_groups` | string | NULL | PILES notation for merging pieces |
| `save_files` | logical | FALSE | Whether to save SVG files |
| `output_dir` | string | "output" | Directory for saved files |

### Type-Specific Parameters (via `...`)

**Hexagonal/Concentric:**

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `tabsize` | numeric | 15 | Tab size in mm |
| `jitter` | numeric | 4 | Randomness in tab shape |
| `do_warp` | logical | TRUE | Apply circular warping |
| `do_trunc` | logical | TRUE | Truncate to boundary |
| `center_shape` | string | "hexagon" | Center shape (concentric only) |

**Voronoi:**

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `relaxation` | integer | 0 | Lloyd relaxation iterations |

**Random:**

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `n_corner` | integer | 5 | Corners per piece (3-8) |

## Return Value

A list with the following components:

| Component | Type | Description |
|-----------|------|-------------|
| `svg_content` | string | Complete SVG document as text |
| `pieces` | list | List of piece objects with paths and metadata |
| `canvas_size` | list | Canvas dimensions (`$width`, `$height`) |
| `files` | list | Paths to saved files (if `save_files = TRUE`) |
| `config` | list | Configuration used to generate the puzzle |

### Piece Object Structure

Each piece in `pieces` contains:

```r
piece <- list(
  id = 1,                    # Piece ID
  path = "M0 0 L100 0...",   # SVG path data
  parsed_segments = list(),   # Parsed path segments
  group_id = 1,              # Fusion group ID
  is_boundary = TRUE         # Whether piece is on boundary
)
```

## Examples

### Basic Usage

```{r}
# Rectangular puzzle
rect <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42
)
cat("Generated", length(rect$pieces), "pieces\n")
```

### Hexagonal Puzzle

```{r}
hex <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE
)
cat("Generated", length(hex$pieces), "hexagonal pieces\n")
```

### With Fusion Groups

```{r}
fused <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42,
  fusion_groups = "1-2,5-6,11-12"
)
# Pieces 1&2, 5&6, and 11&12 are now merged
```

### With Offset

```{r}
separated <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  offset = 10
)
```

### Saving Files

```{r}
#| eval: false
result <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42,
  save_files = TRUE,
  output_dir = "puzzles"
)
cat("Saved to:", result$files$svg, "\n")
```

## Visualization

The result can be visualized using ggpuzzle:

```{r}
#| fig-align: center

ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    diameter = 200,
    seed = 42
  ) +
  scale_fill_viridis_c(guide = "none") +
  coord_fixed() +
  theme_void()
```

Or by writing the SVG:

```{r}
#| eval: false
result <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42
)

# Save as SVG
writeLines(result$svg_content, "puzzle.svg")
```

## See Also

- [PILES Notation](piles-notation.qmd) - Fusion group syntax
- [ggpuzzle Extension](ggpuzzle.qmd) - ggplot2 integration
- [Gallery: Rectangular](../gallery/rectangular.qmd)
- [Gallery: Hexagonal](../gallery/hexagonal.qmd)
