---
title: "Concentric Puzzles"
subtitle: "Rings of pieces radiating from a central shape"
---

```{r}
#| label: setup-concentric
#| include: false
source("quarto/_setup.R")
```

## Overview

Concentric puzzles feature rings of pieces that radiate outward from a central shape. They combine the elegance of hexagonal geometry with a distinctive radial pattern, creating puzzles that are both beautiful and satisfying to assemble.

## Ring Variations

```{r}
#| label: fig-conc-rings
#| fig-align: center
make_conc_rings <- function(r) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = r,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    coord_fixed() +
    theme_void() +
    labs(title = paste0(r, " rings"))
}

make_conc_rings(2) + make_conc_rings(3) + make_conc_rings(4)
```

## Center Shape Options

The center piece can be either a hexagon or a circle:

```{r}
#| label: fig-conc-center-shape
#| layout-ncol: 2
#| fig-cap:
#|   - "Hexagon center"
#|   - "Circle center"

# Hexagon center (default)
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "hexagon",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_void()

# Circle center
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "circle",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_void()
```

## Fill by Ring vs Piece ID

```{r}
#| label: fig-conc-fill-options
#| layout-ncol: 2
#| fig-cap:
#|   - "Fill by ring"
#|   - "Fill by piece_id"

# By ring
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Ring") +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")

# By piece_id
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(piece_id)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Piece") +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")
```

## Color Palettes

```{r}
#| label: fig-conc-palettes
#| fig-align: center

make_conc_palette <- function(pal) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3, seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_void() +
    labs(title = pal)
}

(make_conc_palette("viridis") + make_conc_palette("magma") + make_conc_palette("plasma")) /
(make_conc_palette("inferno") + make_conc_palette("cividis") + make_conc_palette("turbo"))
```

## Offset (Separation)

```{r}
#| label: fig-conc-offset
#| fig-align: center

make_conc_offset <- function(off) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "cividis", guide = "none") +
    coord_fixed() +
    theme_void() +
    labs(title = paste0("offset = ", off))
}

make_conc_offset(0) + make_conc_offset(5) + make_conc_offset(10) + make_conc_offset(15)
```

## Large Concentric Puzzle

```{r}
#| label: fig-conc-large
#| fig-align: center
#| fig-width: 8
#| fig-height: 8

ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 5,
    center_shape = "hexagon",
    seed = 456
  ) +
  scale_fill_viridis_c(
    option = "turbo",
    name = "Ring"
  ) +
  coord_fixed() +
  theme_void() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 16)
  ) +
  labs(title = "5-Ring Concentric Puzzle")
```

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `rings` | integer | 3 | Number of concentric rings |
| `center_shape` | string | "hexagon" | Center shape: "hexagon" or "circle" |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 | Piece separation distance |
| `tabsize` | numeric | 20 | Tab size parameter |
| `jitter` | numeric | 4 | Randomness in tab shape |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## Fill Variables

When using ggplot2 with concentric puzzles, you can map these variables:

| Variable | Description |
|----------|-------------|
| `piece_id` | Unique piece identifier |
| `ring` | Which ring the piece belongs to (0 = center) |
| `is_center` | TRUE if the piece is the center piece |

## Code Example

```{r}
#| label: demo-conc-code
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "concentric",
  grid = c(4),           # 4 rings
  size = c(280),         # 280mm diameter
  seed = 42,
  center_shape = "hexagon",
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "concentric_puzzle.svg")

# Visualize with ggplot2
library(ggplot2)
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 4,
    center_shape = "hexagon",
    seed = 42, offset = 5
  ) +
  scale_fill_viridis_c(option = "plasma") +
  coord_fixed() +
  theme_void()
```
