---
title: "Concentric Puzzles"
subtitle: "Rings of pieces radiating from a central shape"
---

```{r}
#| label: setup-concentric
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Concentric puzzles feature rings of pieces that radiate outward from a central shape. They combine the elegance of hexagonal geometry with a distinctive radial pattern, creating puzzles that are both beautiful and satisfying to assemble.

## Ring Variations

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-rings-ggpuzzle
#| fig-align: center
make_conc_rings <- function(r) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = r,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0(r, " rings"))
}

print(make_conc_rings(2) + make_conc_rings(3) + make_conc_rings(4))
```

## API

```{r}
#| label: fig-conc-rings-api-code
#| eval: false
# API with varying ring counts
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),      # Number of rings
  size = c(200),    # Diameter in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-rings-api
#| echo: false

# Generate ring previews in a grid
rings <- c(2, 3, 4)
items <- lapply(rings, function(r) {
  generate_puzzle(type = "concentric", grid = c(r), size = c(200), seed = 42, fill_palette = "plasma")
})
render_puzzle_grid(items, ncol = 3, labels = paste(rings, "rings"))
```

:::

## Center Shape Options

The center piece can be either a hexagon or a circle:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-center-shape-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "Hexagon center"
#|   - "Circle center"

# Hexagon center (default)
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "hexagon",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Hexagon center"))

# Circle center
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "circle",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Circle center"))
```

## API

```{r}
#| label: fig-conc-center-shape-api
#| layout-ncol: 2
#| fig-cap:
#|   - "Hexagon center"
#|   - "Circle center"

# Hexagon center (default)
result_hex <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  center_shape = "hexagon",
  fill_palette = "magma"
)
render_puzzle_preview(result_hex)

# Circle center
result_circle <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  center_shape = "circle",
  fill_palette = "magma"
)
render_puzzle_preview(result_circle)
```

:::

## Fill by Ring vs Piece ID

```{r}
#| label: fig-conc-fill-options
#| layout-ncol: 2
#| fig-cap:
#|   - "Fill by ring"
#|   - "Fill by piece_id"

# By ring
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Ring") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "By ring") +
  theme(legend.position = "bottom"))

# By piece_id
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(piece_id)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Piece") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "By piece_id") +
  theme(legend.position = "bottom"))
```

## Color Palettes

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-palettes-ggpuzzle
#| fig-align: center

make_conc_palette <- function(pal) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3, seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_conc_palette("viridis") + make_conc_palette("magma") + make_conc_palette("plasma")) /
(make_conc_palette("inferno") + make_conc_palette("cividis") + make_conc_palette("turbo")))
```

## API

```{r}
#| label: fig-conc-palettes-api-code
#| eval: false
# API with fill_palette for colored fills
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  fill_palette = "viridis"  # Options: viridis, magma, plasma, inferno, cividis, turbo
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-palettes-api
#| echo: false

# Generate all palette previews and combine into a grid
palettes <- c("viridis", "magma", "plasma", "inferno", "cividis", "turbo")
items <- lapply(palettes, function(pal) {
  generate_puzzle(type = "concentric", grid = c(3), size = c(200), seed = 42, fill_palette = pal)
})
render_puzzle_grid(items, ncol = 3, labels = palettes)
```

:::

## Fill Direction

Control the spatial order of color assignment. For concentric puzzles, `"reverse"` reverses colors within each ring while keeping the center piece unchanged. This is distinct from `palette_invert` which reverses the dark↔light ends of the palette.

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-direction-ggpuzzle
#| fig-align: center

make_direction_plot <- function(dir) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(fill_order)),
      rings = 3,
      fill_direction = dir,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("fill_direction = \"", dir, "\""))
}

print(make_direction_plot("forward") + make_direction_plot("reverse"))
```

## API

```{r}
#| label: fig-conc-direction-api
#| echo: false

directions <- c("forward", "reverse")
items <- lapply(directions, function(dir) {
  generate_puzzle(type = "concentric", grid = c(3), size = c(200), seed = 42, fill_palette = "viridis", fill_direction = dir)
})
render_puzzle_grid(items, ncol = 2, labels = paste0("fill_direction = \"", directions, "\""))
```

:::

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-offset-ggpuzzle
#| fig-align: center

make_conc_offset <- function(off) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "cividis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_conc_offset(0) + make_conc_offset(5) + make_conc_offset(10) + make_conc_offset(15))
```

## API

```{r}
#| label: fig-conc-offset-api-code
#| eval: false
# API with offset for piece separation
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  offset = 10,  # Separation in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-offset-api
#| echo: false

# Generate offset previews in a grid
offsets <- c(0, 5, 10, 15)
items <- lapply(offsets, function(off) {
  generate_puzzle(type = "concentric", grid = c(3), size = c(200), seed = 42, offset = off, fill_palette = "cividis")
})
render_puzzle_grid(items, ncol = 4, labels = paste("offset =", offsets))
```

:::

## Fusion Groups

Merge pieces together using PILES notation:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-fusion-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "center + ring1"
#|   - "All of ring2"

# Center merged with ring 1
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    fusion_groups = "center-ring1",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "center + ring1"))

# Ring 2 pieces merged
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    fusion_groups = "ring2",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "ring2"))
```

## API

```{r}
#| label: fig-conc-fusion-api
#| layout-ncol: 2
#| fig-cap:
#|   - "center + ring1"
#|   - "All of ring2"

# Center merged with ring 1
result_center_ring1 <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  fusion_groups = "center-ring1",
  fill_palette = "plasma"
)
render_puzzle_preview(result_center_ring1)

# Ring 2 pieces merged
result_ring2 <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  fusion_groups = "ring2",
  fill_palette = "plasma"
)
render_puzzle_preview(result_ring2)
```

:::

## Large Concentric Puzzle

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-large-ggpuzzle
#| fig-align: center
#| fig-width: 8
#| fig-height: 8

ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 5,
    center_shape = "hexagon",
    seed = 456
  ) +
  scale_fill_viridis_c(
    option = "turbo",
    name = "Ring"
  ) +
  coord_fixed() +
  theme_puzzle() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 16)
  ) +
  labs(title = "5-Ring Concentric Puzzle")
```

## API

```{r}
#| label: fig-conc-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "concentric",
  grid = c(5),
  size = c(300),
  seed = 456,
  center_shape = "hexagon",
  fill_palette = "turbo"
)
render_puzzle_preview(result, max_width = "500px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `rings` | integer | 3 | Number of concentric rings |
| `center_shape` | string | "hexagon" | Center shape: "hexagon" or "circle" |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 6 (%) | Tab size as percentage of piece width |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## Fill Variables

When using ggplot2 with concentric puzzles, you can map these variables:

| Variable | Description |
|----------|-------------|
| `piece_id` | Unique piece identifier |
| `ring` | Which ring the piece belongs to (0 = center) |
| `is_center` | TRUE if the piece is the center piece |

## Code Example

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: demo-conc-ggpuzzle
#| eval: false
library(jigsawR)
library(ggplot2)

# Visualize with ggplot2
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 4,
    diameter = 280,
    center_shape = "hexagon",
    seed = 42,
    offset = 5
  ) +
  scale_fill_viridis_c(option = "plasma", name = "Ring") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "4-Ring Concentric Puzzle")
```

## API

```{r}
#| label: demo-conc-api
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "concentric",
  grid = c(4),           # 4 rings
  size = c(280),         # 280mm diameter
  seed = 42,
  center_shape = "hexagon",
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "concentric_puzzle.svg")

# Preview the puzzle
render_puzzle_preview(result)
```

:::

