---
title: "Concentric Puzzles"
subtitle: "Rings of pieces radiating from a central shape"
---

```{r}
#| label: setup-concentric
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Concentric puzzles feature rings of pieces that radiate outward from a central shape. They combine the elegance of hexagonal geometry with a distinctive radial pattern, creating puzzles that are both beautiful and satisfying to assemble.

## Ring Variations

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-rings-ggpuzzle
#| fig-align: center
make_conc_rings <- function(r) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = r,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0(r, " rings"))
}

print(make_conc_rings(2) + make_conc_rings(3) + make_conc_rings(4))
```

## API

```{r}
#| label: fig-conc-rings-api-code
#| eval: false
# API with varying ring counts
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),      # Number of rings
  size = c(200),    # Diameter in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-rings-api
#| results: asis
#| echo: false

# Generate ring previews in a grid
rings <- c(2, 3, 4)

cat('<div style="display: grid; grid-template-columns: repeat(3, 1fr); gap: 1em; margin: 1em 0;">')
for (r in rings) {
  result <- generate_puzzle(
    type = "concentric",
    grid = c(r),
    size = c(200),
    seed = 42
  )
  cat(sprintf('<div style="text-align: center;"><div style="font-weight: bold; margin-bottom: 0.5em;">%d rings</div>%s</div>',
              r, result$svg_content))
}
cat('</div>')
```

:::

## Center Shape Options

The center piece can be either a hexagon or a circle:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-center-shape-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "Hexagon center"
#|   - "Circle center"

# Hexagon center (default)
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "hexagon",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Hexagon center"))

# Circle center
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3,
    center_shape = "circle",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Circle center"))
```

## API

```{r}
#| label: fig-conc-center-shape-api
#| layout-ncol: 2
#| fig-cap:
#|   - "Hexagon center"
#|   - "Circle center"

# Hexagon center (default)
result_hex <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  center_shape = "hexagon"
)
render_puzzle_preview(result_hex)

# Circle center
result_circle <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  center_shape = "circle"
)
render_puzzle_preview(result_circle)
```

:::

## Fill by Ring vs Piece ID

```{r}
#| label: fig-conc-fill-options
#| layout-ncol: 2
#| fig-cap:
#|   - "Fill by ring"
#|   - "Fill by piece_id"

# By ring
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Ring") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "By ring") +
  theme(legend.position = "bottom"))

# By piece_id
print(ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(piece_id)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", name = "Piece") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "By piece_id") +
  theme(legend.position = "bottom"))
```

## Color Palettes

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-palettes-ggpuzzle
#| fig-align: center

make_conc_palette <- function(pal) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3, seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_conc_palette("viridis") + make_conc_palette("magma") + make_conc_palette("plasma")) /
(make_conc_palette("inferno") + make_conc_palette("cividis") + make_conc_palette("turbo")))
```

## API

```{r}
#| label: fig-conc-palettes-api-code
#| eval: false
# API with fill_palette for colored fills
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  seed = 42,
  fill_palette = "viridis"  # Options: viridis, magma, plasma, inferno, cividis, turbo
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-palettes-api
#| results: asis
#| echo: false

# Generate all palette previews and combine into a grid
palettes <- c("viridis", "magma", "plasma", "inferno", "cividis", "turbo")

cat('<div style="display: grid; grid-template-columns: repeat(3, 1fr); gap: 1em; margin: 1em 0;">')
for (pal in palettes) {
  result <- generate_puzzle(
    type = "concentric",
    grid = c(3),
    size = c(200),
    seed = 42,
    fill_palette = pal
  )
  cat(sprintf('<div style="text-align: center;"><div style="font-weight: bold; margin-bottom: 0.5em;">%s</div>%s</div>',
              pal, result$svg_content))
}
cat('</div>')
```

:::

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-offset-ggpuzzle
#| fig-align: center

make_conc_offset <- function(off) {
  ggplot() +
    geom_puzzle_conc(
      aes(fill = after_stat(ring)),
      rings = 3,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "cividis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_conc_offset(0) + make_conc_offset(5) + make_conc_offset(10) + make_conc_offset(15))
```

## API

```{r}
#| label: fig-conc-offset-api-code
#| eval: false
# API with offset for piece separation
result <- generate_puzzle(
  type = "concentric",
  grid = c(3),
  size = c(200),
  offset = 10,  # Separation in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-conc-offset-api
#| results: asis
#| echo: false

# Generate offset previews in a grid
offsets <- c(0, 5, 10, 15)

cat('<div style="display: grid; grid-template-columns: repeat(4, 1fr); gap: 1em; margin: 1em 0;">')
for (off in offsets) {
  result <- generate_puzzle(
    type = "concentric",
    grid = c(3),
    size = c(200),
    seed = 42,
    offset = off
  )
  cat(sprintf('<div style="text-align: center;"><div style="font-weight: bold; margin-bottom: 0.5em;">offset = %d</div>%s</div>',
              off, result$svg_content))
}
cat('</div>')
```

:::

## Large Concentric Puzzle

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-conc-large-ggpuzzle
#| fig-align: center
#| fig-width: 8
#| fig-height: 8

ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 5,
    center_shape = "hexagon",
    seed = 456
  ) +
  scale_fill_viridis_c(
    option = "turbo",
    name = "Ring"
  ) +
  coord_fixed() +
  theme_puzzle() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 16)
  ) +
  labs(title = "5-Ring Concentric Puzzle")
```

## API

```{r}
#| label: fig-conc-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "concentric",
  grid = c(5),
  size = c(300),
  seed = 456,
  center_shape = "hexagon"
)
render_puzzle_preview(result, max_width = "500px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `rings` | integer | 3 | Number of concentric rings |
| `center_shape` | string | "hexagon" | Center shape: "hexagon" or "circle" |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 10 (%) | Tab size as percentage of piece width |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## Fill Variables

When using ggplot2 with concentric puzzles, you can map these variables:

| Variable | Description |
|----------|-------------|
| `piece_id` | Unique piece identifier |
| `ring` | Which ring the piece belongs to (0 = center) |
| `is_center` | TRUE if the piece is the center piece |

## Code Example

::: {.panel-tabset}

## API

```{r}
#| label: demo-conc-api
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "concentric",
  grid = c(4),           # 4 rings
  size = c(280),         # 280mm diameter
  seed = 42,
  center_shape = "hexagon",
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "concentric_puzzle.svg")

# Preview the puzzle
render_puzzle_preview(result)
```

## ggpuzzle

```{r}
#| label: demo-conc-ggpuzzle
#| eval: false
library(jigsawR)
library(ggplot2)

# Visualize with ggplot2
ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 4,
    diameter = 280,
    center_shape = "hexagon",
    seed = 42,
    offset = 5
  ) +
  scale_fill_viridis_c(option = "plasma", name = "Ring") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "4-Ring Concentric Puzzle")
```

:::

