---
title: "Hexagonal Puzzles"
subtitle: "Beautiful honeycomb patterns with curved edges"
---

```{r}
#| label: setup-hexagonal
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Hexagonal puzzles create stunning honeycomb patterns. Each hexagonal piece connects to up to six neighbors through elegantly curved tabs. The result is a visually striking circular puzzle.

## Ring Variations

Hexagonal puzzles are defined by the number of rings:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-rings-ggpuzzle
#| layout-ncol: 3
#| fig-cap:
#|   - "2 rings (7 pieces)"
#|   - "3 rings (19 pieces)"
#|   - "4 rings (37 pieces)"

# 2 rings
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 2, seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "2 rings"))

# 3 rings
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "3 rings"))

# 4 rings
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 4, seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "4 rings"))
```

## API

```{r}
#| label: fig-hex-rings-api
#| layout-ncol: 3
#| fig-cap:
#|   - "2 rings (7 pieces)"
#|   - "3 rings (19 pieces)"
#|   - "4 rings (37 pieces)"

# 2 rings
result_2 <- generate_puzzle(
  type = "hexagonal",
  grid = c(2),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "plasma"
)
render_puzzle_preview(result_2)  # 2 rings

# 3 rings
result_3 <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "plasma"
)
render_puzzle_preview(result_3)  # 3 rings

# 4 rings
result_4 <- generate_puzzle(
  type = "hexagonal",
  grid = c(4),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "plasma"
)
render_puzzle_preview(result_4)  # 4 rings
```

:::

## Piece Count Formula

The number of pieces in a hexagonal puzzle follows the formula:

$$n = 3r(r-1) + 1$$

Where $r$ is the number of rings.

| Rings | Pieces |
|-------|--------|
| 2 | 7 |
| 3 | 19 |
| 4 | 37 |
| 5 | 61 |
| 6 | 91 |

## Warp and Truncation

Hexagonal puzzles support two key transformations:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-warp-trunc-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "warp=TRUE, trunc=TRUE (Default)"
#|   - "warp=TRUE, trunc=FALSE"

# Default: warped and truncated
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    do_warp = TRUE, do_trunc = TRUE,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "warp + trunc"))

# Warped but not truncated (extends beyond circle)
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    do_warp = TRUE, do_trunc = FALSE,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "warp only"))
```

## API

```{r}
#| label: fig-hex-warp-trunc-api
#| layout-ncol: 2
#| fig-cap:
#|   - "warp=TRUE, trunc=TRUE (Default)"
#|   - "warp=TRUE, trunc=FALSE"

# Default: warped and truncated
result_warp_trunc <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "magma"
)
render_puzzle_preview(result_warp_trunc)

# Warped but not truncated
result_warp_only <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = FALSE,
  fill_palette = "magma"
)
render_puzzle_preview(result_warp_only)
```

:::

## Color Palettes

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-palettes-ggpuzzle
#| fig-align: center
make_hex_palette <- function(pal) {
  ggplot() +
    geom_puzzle_hex(
      aes(fill = after_stat(piece_id)),
      rings = 3, seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_hex_palette("viridis") + make_hex_palette("magma") + make_hex_palette("plasma")) /
(make_hex_palette("inferno") + make_hex_palette("cividis") + make_hex_palette("turbo")))
```

## API

```{r}
#| label: fig-hex-palettes-api-code
#| eval: false
# API with fill_palette for colored fills
result <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "viridis"  # Options: viridis, magma, plasma, inferno, cividis, turbo
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-hex-palettes-api
#| echo: false

# Generate all palette previews and combine into a grid
palettes <- c("viridis", "magma", "plasma", "inferno", "cividis", "turbo")
items <- lapply(palettes, function(pal) {
  generate_puzzle(type = "hexagonal", grid = c(3), size = c(200), seed = 42, do_warp = TRUE, do_trunc = TRUE, fill_palette = pal)
})
render_puzzle_grid(items, ncol = 3, labels = palettes)
```

:::

## Fill Direction

Control the spatial order of color assignment. For hexagonal puzzles, `"reverse"` reverses colors within each ring while keeping the center piece unchanged. This is distinct from `palette_invert` which reverses the dark↔light ends of the palette.

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-direction-ggpuzzle
#| fig-align: center

make_direction_plot <- function(dir) {
  ggplot() +
    geom_puzzle_hex(
      aes(fill = after_stat(fill_order)),
      rings = 3,
      fill_direction = dir,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("fill_direction = \"", dir, "\""))
}

print(make_direction_plot("forward") + make_direction_plot("reverse"))
```

## API

```{r}
#| label: fig-hex-direction-api
#| echo: false

directions <- c("forward", "reverse")
items <- lapply(directions, function(dir) {
  generate_puzzle(type = "hexagonal", grid = c(3), size = c(200), seed = 42, do_warp = TRUE, do_trunc = TRUE, fill_palette = "viridis", fill_direction = dir)
})
render_puzzle_grid(items, ncol = 2, labels = paste0("fill_direction = \"", directions, "\""))
```

:::

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-offset-ggpuzzle
#| fig-align: center

make_hex_offset <- function(off) {
  ggplot() +
    geom_puzzle_hex(
      aes(fill = after_stat(piece_id)),
      rings = 3,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_hex_offset(0) + make_hex_offset(5) + make_hex_offset(10) + make_hex_offset(15))
```

## API

```{r}
#| label: fig-hex-offset-api-code
#| eval: false
# API with offset for piece separation
result <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  offset = 10  # Separation in mm
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-hex-offset-api
#| echo: false

# Generate offset previews in a grid
offsets <- c(0, 5, 10, 15)
items <- lapply(offsets, function(off) {
  generate_puzzle(type = "hexagonal", grid = c(3), size = c(200), seed = 42, do_warp = TRUE, do_trunc = TRUE, offset = off, fill_palette = "viridis")
})
render_puzzle_grid(items, ncol = 4, labels = paste("offset =", offsets))
```

:::

## Fusion Groups

Special keywords for hexagonal puzzles:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-fusion-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "center + ring1"
#|   - "All of ring2"

# Center merged with ring 1
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    fusion_groups = "center-ring1",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "center + ring1"))

# Ring 2 pieces merged
print(ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    fusion_groups = "ring2",
    seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "ring2"))
```

## API

```{r}
#| label: fig-hex-fusion-api
#| layout-ncol: 2
#| fig-cap:
#|   - "center + ring1"
#|   - "All of ring2"

# Center merged with ring 1
result_center_ring1 <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fusion_groups = "center-ring1",
  fill_palette = "plasma"
)
render_puzzle_preview(result_center_ring1)

# Ring 2 pieces merged
result_ring2 <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  fusion_groups = "ring2",
  fill_palette = "plasma"
)
render_puzzle_preview(result_ring2)
```

:::

## Large Puzzles

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-hex-large-ggpuzzle
#| fig-align: center
#| fig-width: 8
#| fig-height: 8

ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 5,
    seed = 123
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "5-Ring Hexagonal Puzzle (61 pieces)")
```

## API

```{r}
#| label: fig-hex-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "hexagonal",
  grid = c(5),
  size = c(300),
  seed = 123,
  do_warp = TRUE,
  do_trunc = TRUE,
  fill_palette = "turbo"
)
render_puzzle_preview(result, max_width = "500px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `rings` | integer | 3 | Number of concentric rings |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 6 (%) | Tab size as percentage of edge length |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `do_warp` | logical | FALSE (API) / TRUE (ggpuzzle) | Apply circular warping |
| `do_trunc` | logical | FALSE (API) / TRUE (ggpuzzle) | Truncate to boundary |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## PILES Keywords

Special keywords available for hexagonal puzzles:

| Keyword | Description |
|---------|-------------|
| `center` | The center piece (piece 1) |
| `ring1` | All pieces in ring 1 |
| `ring2` | All pieces in ring 2 |
| `ringN` | All pieces in ring N |
| `boundary` | All outer boundary pieces |

## Code Example

```{r}
#| label: demo-hex-code
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "hexagonal",
  grid = c(4),         # 4 rings
  size = c(250),       # 250mm diameter
  seed = 42,
  do_warp = TRUE,
  do_trunc = TRUE,
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "hexagonal_puzzle.svg")

# Access piece info
n_pieces <- length(result$pieces)
cli::cli_alert_info("Generated {n_pieces} pieces")
```
