---
title: "Random Puzzles"
subtitle: "Unique random shapes with configurable complexity"
---

```{r}
#| label: setup-random
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Random puzzles generate unique, irregular piece shapes using a randomized approach. Unlike the regular patterns of rectangular or hexagonal puzzles, random puzzles create organic, unpredictable boundaries that make each puzzle truly one-of-a-kind.

## Piece Count Variations

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-piece-counts-ggpuzzle
#| fig-align: center
make_rand_pieces <- function(n) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(piece_id)),
      n_interior = n,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "inferno", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0(n, " pieces"))
}

print(make_rand_pieces(5) + make_rand_pieces(10) + make_rand_pieces(20))
```

## API

```{r}
#| label: fig-rand-piece-counts-api-code
#| eval: false
# API with varying piece counts
result <- generate_puzzle(
  type = "random",
  grid = c(10),         # Number of interior points
  size = c(200, 200),   # height, width in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-rand-piece-counts-api
#| echo: false

# Generate piece count previews in a grid
counts <- c(5, 10, 20)
items <- lapply(counts, function(n) {
  generate_puzzle(type = "random", grid = c(n), size = c(200, 200), seed = 42, fill_palette = "inferno")
})
render_puzzle_grid(items, ncol = 3, labels = paste(counts, "pieces"))
```

:::

## Corner Variations

The `n_corner` parameter controls the complexity of each piece's shape:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-corners-ggpuzzle
#| fig-align: center

make_rand_corners <- function(nc, label) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(piece_id)),
      n_interior = 8,
      n_corner = nc,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = label)
}

print(make_rand_corners(3, "3 corners") + make_rand_corners(4, "4 corners") +
make_rand_corners(6, "6 corners") + make_rand_corners(8, "8 corners"))
```

## API

```{r}
#| label: fig-rand-corners-api-code
#| eval: false
# API with n_corner for piece shape complexity
result <- generate_puzzle(
  type = "random",
  grid = c(8),
  size = c(200, 200),
  seed = 42,
  n_corner = 5  # 3-8, higher = more complex shapes
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-rand-corners-api
#| echo: false

# Generate corner variation previews in a grid
corners <- c(3, 4, 6, 8)
items <- lapply(corners, function(nc) {
  generate_puzzle(type = "random", grid = c(8), size = c(200, 200), seed = 42, n_corner = nc, fill_palette = "plasma")
})
render_puzzle_grid(items, ncol = 4, labels = paste(corners, "corners"))
```

:::

## Color Palettes

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-palettes-ggpuzzle
#| fig-align: center

make_rand_palette <- function(pal) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(piece_id)),
      n_interior = 10,
      seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_rand_palette("viridis") + make_rand_palette("magma") + make_rand_palette("plasma")) /
(make_rand_palette("inferno") + make_rand_palette("cividis") + make_rand_palette("turbo")))
```

## API

```{r}
#| label: fig-rand-palettes-api-code
#| eval: false
# API with fill_palette for colored fills
result <- generate_puzzle(
  type = "random",
  grid = c(10),
  size = c(200, 200),
  seed = 42,
  fill_palette = "viridis"  # Options: viridis, magma, plasma, inferno, cividis, turbo
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-rand-palettes-api
#| echo: false

# Generate all palette previews and combine into a grid
palettes <- c("viridis", "magma", "plasma", "inferno", "cividis", "turbo")
items <- lapply(palettes, function(pal) {
  generate_puzzle(type = "random", grid = c(10), size = c(200, 200), seed = 42, fill_palette = pal)
})
render_puzzle_grid(items, ncol = 3, labels = palettes)
```

:::

## Fill Direction

Control the spatial order of color assignment. This is distinct from `palette_invert` which reverses the dark↔light ends of the palette.

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-direction-ggpuzzle
#| fig-align: center

make_direction_plot <- function(dir) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(fill_order)),
      n_interior = 8,
      fill_direction = dir,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("fill_direction = \"", dir, "\""))
}

print(make_direction_plot("forward") + make_direction_plot("reverse"))
```

## API

```{r}
#| label: fig-rand-direction-api
#| echo: false

directions <- c("forward", "reverse")
items <- lapply(directions, function(dir) {
  generate_puzzle(type = "random", grid = c(8), size = c(200, 200), seed = 42, fill_palette = "viridis", fill_direction = dir)
})
render_puzzle_grid(items, ncol = 2, labels = paste0("fill_direction = \"", directions, "\""))
```

:::

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-offset-ggpuzzle
#| fig-align: center

make_rand_offset <- function(off) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(piece_id)),
      n_interior = 9,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "cividis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_rand_offset(0) + make_rand_offset(5) + make_rand_offset(10) + make_rand_offset(20))
```

## API

```{r}
#| label: fig-rand-offset-api-code
#| eval: false
# API with offset for piece separation
result <- generate_puzzle(
  type = "random",
  grid = c(9),
  size = c(200, 200),
  offset = 10,  # Separation in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-rand-offset-api
#| echo: false

# Generate offset previews in a grid
offsets <- c(0, 5, 10, 20)
items <- lapply(offsets, function(off) {
  generate_puzzle(type = "random", grid = c(9), size = c(200, 200), seed = 42, offset = off, fill_palette = "cividis")
})
render_puzzle_grid(items, ncol = 4, labels = paste("offset =", offsets))
```

:::

## Seed Variations

Each seed creates a completely unique puzzle:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-seeds-ggpuzzle
#| fig-align: center

make_rand_seed <- function(s) {
  ggplot() +
    geom_puzzle_random(
      aes(fill = after_stat(piece_id)),
      n_interior = 8,
      seed = s
    ) +
    scale_fill_viridis_c(option = "turbo", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("seed = ", s))
}

print(make_rand_seed(1) + make_rand_seed(42) + make_rand_seed(123) + make_rand_seed(999))
```

## API

```{r}
#| label: fig-rand-seeds-api-code
#| eval: false
# API with different seeds for unique patterns
result <- generate_puzzle(
  type = "random",
  grid = c(8),
  size = c(200, 200),
  seed = 42  # Change seed for different patterns
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-rand-seeds-api
#| echo: false

# Generate seed variation previews in a grid
seeds <- c(1, 42, 123, 999)
items <- lapply(seeds, function(s) {
  generate_puzzle(type = "random", grid = c(8), size = c(200, 200), seed = s, fill_palette = "turbo")
})
render_puzzle_grid(items, ncol = 4, labels = paste("seed =", seeds))
```

:::

::: {.callout-tip}
## Aspect Ratios
Random puzzles handle any aspect ratio - wide, square, or tall. See the [Aspect Ratios Tutorial](../tutorials/aspect-ratios.qmd) for examples comparing all puzzle types.
:::

## Fusion Groups

Merge pieces together using PILES notation:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-fusion-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 12-piece"
#|   - "Adjacent pieces fused"

# Standard
print(ggplot() +
  geom_puzzle_random(
    aes(fill = after_stat(piece_id)),
    n_interior = 6,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Standard"))

# Adjacent pieces fused
print(ggplot() +
  geom_puzzle_random(
    aes(fill = after_stat(piece_id)),
    n_interior = 6,
    seed = 42,
    fusion_groups = "1-2-3,5-6"
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Pieces fused"))
```

## API

```{r}
#| label: fig-rand-fusion-api
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 12-piece"
#|   - "Adjacent pieces fused"

# Standard
result_standard <- generate_puzzle(
  type = "random",
  grid = c(6),
  size = c(200, 200),
  seed = 42,
  fill_palette = "magma"
)
render_puzzle_preview(result_standard, title = "Standard")

# Adjacent pieces fused
result_fused <- generate_puzzle(
  type = "random",
  grid = c(6),
  size = c(200, 200),
  seed = 42,
  fill_palette = "magma",
  fusion_groups = "1-2-3,5-6"
)
render_puzzle_preview(result_fused, title = "Pieces fused")
```

:::

## Large Random Puzzle

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rand-large-ggpuzzle
#| fig-align: center
#| fig-width: 8
#| fig-height: 6

ggplot() +
  geom_puzzle_random(
    aes(fill = after_stat(piece_id)),
    n_interior = 30,
    width = 400, height = 300,
    seed = 456
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "30-Piece Random Puzzle")
```

## API

```{r}
#| label: fig-rand-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "random",
  grid = c(30),
  size = c(300, 400),
  seed = 456,
  fill_palette = "turbo"
)
render_puzzle_preview(result, max_width = "600px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `n_interior` | integer | 12 | Number of interior points (influences piece count) |
| `n_corner` | integer | 4 | Corners per piece (3-8) |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 6 (%) | Tab size as percentage of edge length |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | auto (random) / Inf (others) | Maximum tab size in mm. Auto-calculated for random puzzles to prevent overlapping tabs |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## Corner Complexity Guide

| n_corner | Shape | Complexity |
|----------|-------|------------|
| 3 | Triangular | Simple, angular |
| 4 | Quadrilateral | Classic, balanced |
| 5 | Pentagonal | Natural, organic |
| 6 | Hexagonal | Honeycomb-like |
| 7-8 | Complex | Highly irregular |

## When to Use Random Puzzles

Random puzzles are perfect for:

- **Maximum uniqueness**: No two puzzles alike
- **Artistic applications**: Abstract, modern designs
- **Challenge modes**: Irregular shapes are harder to assemble
- **Custom piece counts**: Exact control over difficulty

## Code Example

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: demo-rand-ggpuzzle
#| fig-align: center

ggplot() +
  geom_puzzle_random(
    aes(fill = after_stat(piece_id)),
    n_interior = 15,
    width = 400, height = 300,  # width=400mm, height=300mm
    n_corner = 5,
    seed = 42, offset = 5
  ) +
  scale_fill_viridis_c(option = "inferno", guide = "none") +
  coord_fixed() +
  theme_puzzle()
```

## API

```{r}
#| label: demo-rand-api
#| fig-align: center

result <- generate_puzzle(
  type = "random",
  grid = c(15),          # 15 pieces
  size = c(300, 400),    # height=300mm, width=400mm
  seed = 42,
  n_corner = 5,          # Pentagonal pieces
  offset = 5
)

render_puzzle_preview(result)
```

:::

::: {.callout-note}
## Saving to File
The API approach returns a result object that can be saved to SVG:
```r
writeLines(result$svg_content, "random_puzzle.svg")
```
:::
