---
title: "Rectangular Puzzles"
subtitle: "Classic jigsaw puzzles with interlocking tabs"
---

```{r}
#| label: setup-rectangular
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Rectangular puzzles are the classic jigsaw format with interlocking tabs on all sides. Each piece connects to its neighbors through curved tabs that can point inward or outward.

## Size Variations

Explore different grid sizes:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rect-sizes-ggpuzzle
#| layout-ncol: 3
#| fig-cap:
#|   - "2x2 (4 pieces)"
#|   - "4x3 (12 pieces)"
#|   - "6x5 (30 pieces)"

print(ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 2, rows = 2,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "2x2"))

print(ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 4, rows = 3,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "4x3"))

print(ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 6, rows = 5,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "6x5"))
```

## API

```{r}
#| label: fig-rect-sizes-api
#| layout-ncol: 3
#| fig-cap:
#|   - "2x2 (4 pieces)"
#|   - "4x3 (12 pieces)"
#|   - "6x5 (30 pieces)"

result_2x2 <- generate_puzzle(
  type = "rectangular",
  grid = c(2, 2),
  size = c(200, 200),
  seed = 42
)
render_puzzle_preview(result_2x2, title = "2x2")

result_4x3 <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42
)
render_puzzle_preview(result_4x3, title = "4x3")

result_6x5 <- generate_puzzle(
  type = "rectangular",
  grid = c(6, 5),
  size = c(600, 500),
  seed = 42
)
render_puzzle_preview(result_6x5, title = "6x5")
```

:::

## Color Palettes

All viridis palettes work beautifully with puzzle pieces:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rect-palettes-ggpuzzle
#| fig-align: center
make_palette_plot <- function(pal) {
  ggplot() +
    geom_puzzle_rect(
      aes(fill = after_stat(piece_id)),
      cols = 4, rows = 3,
      seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_palette_plot("viridis") + make_palette_plot("magma") + make_palette_plot("plasma")) /
(make_palette_plot("inferno") + make_palette_plot("cividis") + make_palette_plot("turbo")))
```

## API

```{r}
#| label: fig-rect-palettes-api
#| fig-align: center
make_palette_plot_api <- function(pal) {
  result <- generate_puzzle(
    type = "rectangular",
    grid = c(4, 3),
    size = c(400, 300),
    seed = 42
  )
  render_puzzle_preview(result, title = pal, palette = pal)
}

print((make_palette_plot_api("viridis") + make_palette_plot_api("magma") + make_palette_plot_api("plasma")) /
(make_palette_plot_api("inferno") + make_palette_plot_api("cividis") + make_palette_plot_api("turbo")))
```

:::

## Offset (Separation)

Control spacing between pieces:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rect-offset-ggpuzzle
#| fig-align: center

make_offset_plot <- function(off) {
  ggplot() +
    geom_puzzle_rect(
      aes(fill = after_stat(piece_id)),
      cols = 3, rows = 3,
      offset = off,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "plasma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_offset_plot(0) + make_offset_plot(5) + make_offset_plot(10) + make_offset_plot(20))
```

## API

```{r}
#| label: fig-rect-offset-api
#| fig-align: center

make_offset_plot_api <- function(off) {
  result <- generate_puzzle(
    type = "rectangular",
    grid = c(3, 3),
    size = c(300, 300),
    offset = off,
    seed = 42
  )
  render_puzzle_preview(result, title = paste0("offset = ", off))
}

print(make_offset_plot_api(0) + make_offset_plot_api(5) + make_offset_plot_api(10) + make_offset_plot_api(20))
```

:::

## Fusion Groups

Merge pieces together using PILES notation:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-rect-fusion-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 4x3"
#|   - "Row fusion: R1,R2,R3"

# Standard
print(ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 4, rows = 3,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Standard"))

# With row fusion
print(ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 4, rows = 3,
    seed = 42,
    fusion_groups = "R1,R2,R3"
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Row fusion"))
```

## API

```{r}
#| label: fig-rect-fusion-api
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 4x3"
#|   - "Row fusion: R1,R2,R3"

# Standard
result_standard <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42
)
render_puzzle_preview(result_standard, title = "Standard")

# With row fusion
result_fusion <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42,
  fusion_groups = "R1,R2,R3"
)
render_puzzle_preview(result_fusion, title = "Row fusion")
```

:::

::: {.callout-tip}
## Aspect Ratios
Rectangular puzzles handle any aspect ratio - wide, square, or tall. See the [Aspect Ratios Tutorial](../tutorials/aspect-ratios.qmd) for examples comparing all puzzle types.
:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `cols` | integer | 3 | Number of columns |
| `rows` | integer | 3 | Number of rows |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 10 (%) | Tab size as percentage of piece width |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |
| `fusion_style` | string | "none" | Style for fused edges: "none", "dashed", "solid" |

## Code Example

::: {.panel-tabset}

## API

```{r}
#| label: demo-rect-usage-api
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "rectangular",
  grid = c(6, 4),      # 6 columns, 4 rows
  size = c(600, 400),  # 600mm x 400mm
  seed = 42,
  offset = 5           # Slight separation
)

# Save SVG
writeLines(result$svg_content, "rectangular_puzzle.svg")
```

## ggpuzzle

```{r}
#| label: demo-rect-usage-ggpuzzle
#| eval: false
library(jigsawR)
library(ggplot2)

# Visualize with ggplot2
ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 6, rows = 4,
    seed = 42, offset = 5
  ) +
  scale_fill_viridis_c() +
  coord_fixed() +
  theme_puzzle()

# Save as PNG
ggsave("rectangular_puzzle.png", width = 10, height = 8, dpi = 300)
```

:::
