---
title: "Voronoi Puzzles"
subtitle: "Organic, natural-looking tessellations"
---

```{r}
#| label: setup-voronoi
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Voronoi puzzles create organic, natural-looking pieces based on Voronoi tessellation. Each piece is a convex polygon formed by the regions closest to a set of seed points. The result resembles patterns found in nature, like honeycomb cells, giraffe spots, or cracked mud.

## Cell Count Variations

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-cell-counts-ggpuzzle
#| fig-align: center
make_vor_cells <- function(n) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = n,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "turbo", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0(n, " cells"))
}

print(make_vor_cells(10) + make_vor_cells(25) + make_vor_cells(50))
```

## API

```{r}
#| label: fig-vor-cell-counts-api-code
#| eval: false
# API with varying cell counts
result <- generate_puzzle(
  type = "voronoi",
  grid = c(25),         # Number of cells
  size = c(200, 200),   # height, width in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-vor-cell-counts-api
#| echo: false

# Generate cell count previews in a grid
cells <- c(10, 25, 50)
items <- lapply(cells, function(n) {
  generate_puzzle(type = "voronoi", grid = c(n), size = c(200, 200), seed = 42, fill_palette = "turbo")
})
render_puzzle_grid(items, ncol = 3, labels = paste(cells, "cells"))
```

:::

::: {.callout-tip}
## Aspect Ratios
Voronoi puzzles adapt to any aspect ratio - wide, square, or tall. See the [Aspect Ratios Tutorial](../tutorials/aspect-ratios.qmd) for examples comparing all puzzle types.
:::

## Color Palettes

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-palettes-ggpuzzle
#| fig-align: center

make_vor_palette <- function(pal) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 20,
      seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_vor_palette("viridis") + make_vor_palette("magma") + make_vor_palette("plasma")) /
(make_vor_palette("inferno") + make_vor_palette("cividis") + make_vor_palette("turbo")))
```

## API

```{r}
#| label: fig-vor-palettes-api-code
#| eval: false
# API with fill_palette for colored fills
result <- generate_puzzle(
  type = "voronoi",
  grid = c(20),
  size = c(200, 200),
  seed = 42,
  fill_palette = "viridis"  # Options: viridis, magma, plasma, inferno, cividis, turbo
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-vor-palettes-api
#| echo: false

# Generate all palette previews and combine into a grid
palettes <- c("viridis", "magma", "plasma", "inferno", "cividis", "turbo")
items <- lapply(palettes, function(pal) {
  generate_puzzle(type = "voronoi", grid = c(20), size = c(200, 200), seed = 42, fill_palette = pal)
})
render_puzzle_grid(items, ncol = 3, labels = palettes)
```

:::

## Fill Direction

Control the spatial order of color assignment. This is distinct from `palette_invert` which reverses the dark↔light ends of the palette.

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-direction-ggpuzzle
#| fig-align: center

make_direction_plot <- function(dir) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(fill_order)),
      n_cells = 20,
      fill_direction = dir,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("fill_direction = \"", dir, "\""))
}

print(make_direction_plot("forward") + make_direction_plot("reverse"))
```

## API

```{r}
#| label: fig-vor-direction-api
#| echo: false

directions <- c("forward", "reverse")
items <- lapply(directions, function(dir) {
  generate_puzzle(type = "voronoi", grid = c(20), size = c(200, 200), seed = 42, fill_palette = "viridis", fill_direction = dir)
})
render_puzzle_grid(items, ncol = 2, labels = paste0("fill_direction = \"", directions, "\""))
```

:::

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-offset-ggpuzzle
#| fig-align: center

make_vor_offset <- function(off) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 20,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_vor_offset(0) + make_vor_offset(5) + make_vor_offset(10) + make_vor_offset(20))
```

## API

```{r}
#| label: fig-vor-offset-api-code
#| eval: false
# API with offset for piece separation
result <- generate_puzzle(
  type = "voronoi",
  grid = c(20),
  size = c(200, 200),
  offset = 10,  # Separation in mm
  seed = 42
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-vor-offset-api
#| echo: false

# Generate offset previews in a grid
offsets <- c(0, 5, 10, 20)
items <- lapply(offsets, function(off) {
  generate_puzzle(type = "voronoi", grid = c(20), size = c(200, 200), seed = 42, offset = off, fill_palette = "viridis")
})
render_puzzle_grid(items, ncol = 4, labels = paste("offset =", offsets))
```

:::

## Seed Variations

Different seeds create completely different tessellations:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-seeds-ggpuzzle
#| fig-align: center

make_vor_seed <- function(s) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 15,
      seed = s
    ) +
    scale_fill_viridis_c(option = "magma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("seed = ", s))
}

print(make_vor_seed(1) + make_vor_seed(42) + make_vor_seed(123) + make_vor_seed(999))
```

## API

```{r}
#| label: fig-vor-seeds-api-code
#| eval: false
# API with different seeds for unique patterns
result <- generate_puzzle(
  type = "voronoi",
  grid = c(15),
  size = c(200, 200),
  seed = 42  # Change seed for different patterns
)
render_puzzle_preview(result)
```

```{r}
#| label: fig-vor-seeds-api
#| echo: false

# Generate seed variation previews in a grid
seeds <- c(1, 42, 123, 999)
items <- lapply(seeds, function(s) {
  generate_puzzle(type = "voronoi", grid = c(15), size = c(200, 200), seed = s, fill_palette = "magma")
})
render_puzzle_grid(items, ncol = 4, labels = paste("seed =", seeds))
```

:::

## Fusion Groups

Merge pieces together using PILES notation:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-fusion-ggpuzzle
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 12-cell"
#|   - "Adjacent pieces fused"

# Standard
print(ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 12,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Standard"))

# Adjacent pieces fused
print(ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 12,
    seed = 42,
    fusion_groups = "1-2-3,5-6"
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "Pieces fused"))
```

## API

```{r}
#| label: fig-vor-fusion-api
#| layout-ncol: 2
#| fig-cap:
#|   - "Standard 12-cell"
#|   - "Adjacent pieces fused"

# Standard
result_standard <- generate_puzzle(
  type = "voronoi",
  grid = c(12),
  size = c(200, 200),
  seed = 42,
  fill_palette = "magma"
)
render_puzzle_preview(result_standard, title = "Standard")

# Adjacent pieces fused
result_fused <- generate_puzzle(
  type = "voronoi",
  grid = c(12),
  size = c(200, 200),
  seed = 42,
  fill_palette = "magma",
  fusion_groups = "1-2-3,5-6"
)
render_puzzle_preview(result_fused, title = "Pieces fused")
```

:::

## Large Voronoi Puzzle

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-large-ggpuzzle
#| fig-align: center
#| fig-width: 8
#| fig-height: 6

ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 80,
    width = 400, height = 300,
    seed = 789
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "80-Cell Voronoi Puzzle")
```

## API

```{r}
#| label: fig-vor-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "voronoi",
  grid = c(80),
  size = c(300, 400),
  seed = 789,
  fill_palette = "turbo"
)
render_puzzle_preview(result, max_width = "600px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `n_cells` | integer | 12 | Number of Voronoi cells (pieces) |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 6 (%) | Tab size as percentage of piece width |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `point_distribution` | string | "fermat" | Seed point distribution method |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## About Voronoi Tessellation

Voronoi tessellation divides a plane into regions based on distance to seed points. Each region consists of all points closer to its seed than to any other seed. This creates:

- **Convex polygons**: Each piece is a convex polygon
- **Natural appearance**: Resembles patterns in nature
- **Variable sizes**: Piece sizes depend on seed point distribution

## Code Example

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: demo-vor-code-ggpuzzle
#| eval: false
library(jigsawR)
library(ggplot2)

ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 35,
    width = 400, height = 300,
    seed = 42,
    offset = 5
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle()
```

## API

```{r}
#| label: demo-vor-code-api
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "voronoi",
  grid = c(35),          # 35 cells
  size = c(300, 400),    # height=300mm, width=400mm
  seed = 42,
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "voronoi_puzzle.svg")

# Preview
render_puzzle_preview(result)
```

:::

## When to Use Voronoi

Voronoi puzzles are ideal for:

- **Organic themes**: Nature, landscapes, biological subjects
- **Artistic effects**: Modern art, abstract designs
- **Variable difficulty**: More cells = harder puzzle
- **Unique designs**: Every seed creates a different pattern
