---
title: "Voronoi Puzzles"
subtitle: "Organic, natural-looking tessellations"
---

```{r}
#| label: setup-voronoi
#| include: false
# Flexible path: works from project root (Quarto) or quarto/gallery/ (RStudio interactive)
setup_path <- if (file.exists("quarto/_setup.R")) "quarto/_setup.R" else "../_setup.R"
source(setup_path)
```

## Overview

Voronoi puzzles create organic, natural-looking pieces based on Voronoi tessellation. Each piece is a convex polygon formed by the regions closest to a set of seed points. The result resembles patterns found in nature, like honeycomb cells, giraffe spots, or cracked mud.

## Cell Count Variations

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-cell-counts-ggpuzzle
#| fig-align: center
make_vor_cells <- function(n) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = n,
      seed = 42
    ) +
    scale_fill_viridis_c(option = "turbo", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0(n, " cells"))
}

print(make_vor_cells(10) + make_vor_cells(25) + make_vor_cells(50))
```

## API

```{r}
#| label: fig-vor-cell-counts-api
#| layout-ncol: 3
#| fig-cap:
#|   - "10 cells"
#|   - "25 cells"
#|   - "50 cells"

for (n in c(10, 25, 50)) {
  result <- generate_puzzle(
    type = "voronoi",
    grid = c(n),
    size = c(200, 200),
    seed = 42
  )
  render_puzzle_preview(result)
}
```

:::

::: {.callout-tip}
## Aspect Ratios
Voronoi puzzles adapt to any aspect ratio - wide, square, or tall. See the [Aspect Ratios Tutorial](../tutorials/aspect-ratios.qmd) for examples comparing all puzzle types.
:::

## Color Palettes

```{r}
#| label: fig-vor-color-palettes
#| fig-align: center

make_vor_palette <- function(pal) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 20,
      seed = 42
    ) +
    scale_fill_viridis_c(option = pal, guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = pal)
}

print((make_vor_palette("viridis") + make_vor_palette("magma") + make_vor_palette("plasma")) /
(make_vor_palette("inferno") + make_vor_palette("cividis") + make_vor_palette("turbo")))
```

## Offset (Separation)

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-offset-ggpuzzle
#| fig-align: center

make_vor_offset <- function(off) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 20,
      offset = off, seed = 42
    ) +
    scale_fill_viridis_c(option = "viridis", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("offset = ", off))
}

print(make_vor_offset(0) + make_vor_offset(5) + make_vor_offset(10) + make_vor_offset(20))
```

## API

```{r}
#| label: fig-vor-offset-api
#| layout-ncol: 4
#| fig-cap:
#|   - "offset = 0"
#|   - "offset = 5"
#|   - "offset = 10"
#|   - "offset = 20"

for (off in c(0, 5, 10, 20)) {
  result <- generate_puzzle(
    type = "voronoi",
    grid = c(20),
    size = c(200, 200),
    seed = 42,
    offset = off
  )
  render_puzzle_preview(result)
}
```

:::

## Seed Variations

Different seeds create completely different tessellations:

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-seeds-ggpuzzle
#| fig-align: center

make_vor_seed <- function(s) {
  ggplot() +
    geom_puzzle_voronoi(
      aes(fill = after_stat(piece_id)),
      n_cells = 15,
      seed = s
    ) +
    scale_fill_viridis_c(option = "magma", guide = "none") +
    coord_fixed() +
    theme_puzzle() +
    labs(title = paste0("seed = ", s))
}

print(make_vor_seed(1) + make_vor_seed(42) + make_vor_seed(123) + make_vor_seed(999))
```

## API

```{r}
#| label: fig-vor-seeds-api
#| layout-ncol: 4
#| fig-cap:
#|   - "seed = 1"
#|   - "seed = 42"
#|   - "seed = 123"
#|   - "seed = 999"

for (s in c(1, 42, 123, 999)) {
  result <- generate_puzzle(
    type = "voronoi",
    grid = c(15),
    size = c(200, 200),
    seed = s
  )
  render_puzzle_preview(result)
}
```

:::

## Large Voronoi Puzzle

::: {.panel-tabset}

## ggpuzzle

```{r}
#| label: fig-vor-large-ggpuzzle
#| fig-align: center
#| fig-width: 10
#| fig-height: 6

ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 80,
    seed = 789
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle() +
  labs(title = "80-Cell Voronoi Puzzle")
```

## API

```{r}
#| label: fig-vor-large-api
#| fig-align: center

result <- generate_puzzle(
  type = "voronoi",
  grid = c(80),
  size = c(400, 300),
  seed = 789
)
render_puzzle_preview(result, max_width = "600px")
```

:::

## Parameters Reference

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `n_cells` | integer | 12 | Number of Voronoi cells (pieces) |
| `seed` | integer | random | Random seed for reproducibility |
| `offset` | numeric | 0 (mm) | Piece separation distance |
| `tabsize` | numeric | 10 (%) | Tab size as percentage of piece width |
| `jitter` | numeric | 2 (%) | Randomness in tab shape as percentage |
| `min_tab_size` | numeric | 0 (mm) | Minimum tab size in millimeters |
| `max_tab_size` | numeric | Inf (mm) | Maximum tab size in millimeters |
| `point_distribution` | string | "fermat" | Seed point distribution method |
| `fusion_groups` | string | NULL | PILES notation for fusing pieces |

## About Voronoi Tessellation

Voronoi tessellation divides a plane into regions based on distance to seed points. Each region consists of all points closer to its seed than to any other seed. This creates:

- **Convex polygons**: Each piece is a convex polygon
- **Natural appearance**: Resembles patterns in nature
- **Variable sizes**: Piece sizes depend on seed point distribution

### Lloyd Relaxation

Applying Lloyd relaxation (setting `relaxation > 0`) moves seed points toward the center of their cells, creating more uniform piece sizes:

```{r}
#| label: fig-vor-relaxation
#| layout-ncol: 2
#| fig-cap:
#|   - "No relaxation"
#|   - "With relaxation (10 iterations)"
#| eval: false

# Note: relaxation parameter may need implementation verification
# These examples show the conceptual difference
```

## Code Example

::: {.panel-tabset}

## API

```{r}
#| label: demo-vor-code-api
#| eval: false
library(jigsawR)

# Generate puzzle
result <- generate_puzzle(
  type = "voronoi",
  grid = c(35),          # 35 cells
  size = c(400, 300),    # 400mm x 300mm
  seed = 42,
  offset = 5
)

# Save SVG
writeLines(result$svg_content, "voronoi_puzzle.svg")

# Preview
render_puzzle_preview(result)
```

## ggpuzzle

```{r}
#| label: demo-vor-code-ggpuzzle
#| eval: false
library(jigsawR)
library(ggplot2)

ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 35,
    width = 400, height = 300,
    seed = 42,
    offset = 5
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_puzzle()
```

:::

## When to Use Voronoi

Voronoi puzzles are ideal for:

- **Organic themes**: Nature, landscapes, biological subjects
- **Artistic effects**: Modern art, abstract designs
- **Variable difficulty**: More cells = harder puzzle
- **Unique designs**: Every seed creates a different pattern
