---
title: "jigsawR"
subtitle: "Generate Beautiful Jigsaw Puzzles in R"
page-layout: full
---

::: hero-section
# jigsawR {.hero-title}
:::

::: hero-section
# Generate stunning, mathematically precise jigsaw puzzles with ease. Five puzzle types, infinite possibilities. {.hero-subtitle}
:::

```{r}
#| label: setup
#| include: false
source("quarto/_setup.R")
```

```{r}
#| label: hero-puzzle
#| echo: false
#| fig-align: center
#| out-width: "100%"

# Create a visually appealing hexagonal puzzle
set.seed(42)
p <- ggplot() +

  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = NA)
  )
print(p)
```

## Installation

``` r
# Install from GitHub
devtools::install_github("pjt222/jigsawR")
```

::::::::::::::::::::: feature-grid
::::: feature-card
::: feature-icon
🧩
:::

::: feature-title
5 Puzzle Types
:::

Rectangular, Hexagonal, Concentric, Voronoi, and Random shapes
:::::

::::: feature-card
::: feature-icon
🎨
:::

::: feature-title
Beautiful Palettes
:::

Full viridis color palette support with gradient fills
:::::

::::: feature-card
::: feature-icon
📊
:::

::: feature-title
ggplot2 Integration
:::

Native geom_puzzle\_\*() functions for data visualization
:::::

::::: feature-card
::: feature-icon
🔗
:::

::: feature-title
Fusion Groups
:::

Merge pieces using PILES notation for custom layouts
:::::

::::: feature-card
::: feature-icon
⚡
:::

::: feature-title
High Performance
:::

C++ backend with Rcpp for fast puzzle generation
:::::

::::: feature-card
::: feature-icon
📱
:::

::: feature-title
Shiny App
:::

Interactive web app for puzzle creation and download
:::::
:::::::::::::::::::::

## Quick Start

```{r}
#| label: quick-start
#| eval: false
library(jigsawR)

# Generate a 4x3 rectangular puzzle
result <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  seed = 42
)

# Save as SVG
writeLines(result$svg_content, "my_puzzle.svg")
```

## Gallery Preview

::: puzzle-gallery
```{r}
#| label: gallery-preview
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
p1 <- ggplot() +
  geom_puzzle_rect(
    aes(fill = after_stat(piece_id)),
    cols = 3, rows = 3,
    seed = 42
  ) +
  scale_fill_viridis_c(option = "viridis", guide = "none") +
  coord_fixed() +
  theme_void() +
  labs(title = "Rectangular")

p2 <- ggplot() +
  geom_puzzle_hex(
    aes(fill = after_stat(piece_id)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "magma", guide = "none") +
  coord_fixed() +
  theme_void() +
  labs(title = "Hexagonal")

p3 <- ggplot() +
  geom_puzzle_conc(
    aes(fill = after_stat(ring)),
    rings = 3, seed = 42
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +
  coord_fixed() +
  theme_void() +
  labs(title = "Concentric")

p4 <- ggplot() +
  geom_puzzle_voronoi(
    aes(fill = after_stat(piece_id)),
    n_cells = 15, seed = 42,
    tabsize = 15, jitter = 3
  ) +
  scale_fill_viridis_c(option = "turbo", guide = "none") +
  coord_fixed() +
  theme_void() +
  labs(title = "Voronoi")

print(p1 + p2 + p3 + p4)
```
:::

[Explore the Gallery →](gallery/rectangular.qmd){.btn .btn-primary}

## ggplot2 Integration

Map data to puzzle pieces for creative data visualization:

```{r}
#| label: ggplot2-integration
#| fig-align: center
#| out-width: "60%"

# Create data
data <- data.frame(
  region = 1:7,
  value = c(23, 45, 12, 67, 34, 89, 56)
)

# Visualize with puzzle pieces
ggplot(data) +
  geom_puzzle_hex(
    aes(fill = value),
    rings = 2,
    seed = 123
  ) +
  scale_fill_viridis_c(
    name = "Value",
    option = "plasma"
  ) +
  coord_fixed() +
  theme_minimal() +
  labs(title = "Data as Puzzle Pieces")
```

[Learn More About ggpuzzle →](api/ggpuzzle.qmd){.btn .btn-outline-primary}

## PILES Notation

Powerful piece fusion with intuitive syntax:

```{r}
#| label: piles-demo
#| eval: false

# Merge center with ring 1
result <- generate_puzzle(
  type = "hexagonal",
  grid = c(3),
  size = c(200),
  fusion_groups = "center-ring1"
)

# Custom fusion groups
result <- generate_puzzle(
  type = "rectangular",
  grid = c(4, 3),
  size = c(400, 300),
  fusion_groups = "1-2-3,7-8-9,10-11-12"
)
```

[PILES Notation Guide →](api/piles-notation.qmd){.btn .btn-outline-primary}

------------------------------------------------------------------------

::: text-center
**Ready to create puzzles?**

[Get Started](getting-started.qmd){.btn .btn-lg .btn-primary} [View on GitHub](https://github.com/pjt222/jigsawR){.btn .btn-lg .btn-outline-secondary}
:::
